---
title: "How income varies across different categories of college majors"
author: "Jonathan Sinclair"
date: "07/04/2021"
output:
    html_document:
      keep_md: true 
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


<br/><br/><br/><br/>


##### 0 - Load Libraries


```{r}
devtools::install_github("jhudsl/collegeIncome")
devtools::install_github("jhudsl/matahari")

```


```{r, include=FALSE, message=FALSE, warning=FALSE}
library(collegeIncome)
library(matahari)
dance_start(value = FALSE,
            contents = FALSE)
dance_save("~/Desktop/college_major_analysis.rds")
```


```{r}
library(tidyverse)
library(janitor)
library(GGally)
```

<br/><br/><br/><br/>





##### 1 - Load Data
```{r}
data(college)
```

<br/><br/><br/><br/>





##### 2 - Descriptive Statistics
```{r}
dim(college)
glimpse(college)
```

**How many majors / major categories do we have?**

```{r}
college %>% select(major) %>% n_distinct()
college %>% select(major_category) %>% n_distinct()

sort(table(college$major_category))
```

We can see that we have

* 173 Majors
* 15 Major Categories
  * which range from having 1, to 29 participants


```{r}
#look at the average median incomes by major_category
average_median_incomes <- college %>%
  select(major_category, median) %>%
  group_by(major_category) %>%
  summarise(median_mean = mean(median, na.rm = TRUE)) %>%
  arrange(desc(median_mean))

average_median_incomes
```

We can see by looking at the average of median incomes (by major category) that higher median incomes are associated with

1. Business
2. Biology & Life Science
3. Agriculture & Natural resources

<br/><br/><br/><br/>





##### 3 - EDA

We will now use a box-plot to look at the distribution of of incomes for each major_category

```{r}
g <- ggplot(college, aes(x=major_category,
                         y=median / 1000,
                         color = major_category))
g <- g + geom_boxplot()
g <- g + coord_flip()
g <- g + ylab("Median income (Thousands of dollars)")
g <- g + xlab("Major Category")
g <- g + theme(legend.position = "none")
g
```

The distribution of median incomes doesn't show any discernable pattern. But we can see that **Business** appears to be highest overall.

<br/><br/><br/><br/>





##### 4 - Modelling

Here, we generate a Regression Model, with `major category` regressed

```{r}
fit <- lm(median ~ major_category, data=college)
round(summary(fit)$coef, 2)
```

The model which regresses median income onto major categories, is showing *Agriculture & Natural Resources* as an intercept (our referenced major category). 

The other coefficients show the difference of the mean from that category, to our reference category. 

The p-value of the coefficients is the probability of a t-test where the mean of each coefficient is compared with the mean of the reference variable.

**Model Interpretation**

* Mean of median income for students of *Argriculture & Natural Resources* $43,500
* Difference in means (of median income) between *Argiculture & Natrual Resources* and *Business* is $5,653.85.
  * This is the highest largest mean difference of all major categories.
  

**Checking Model validity**

```{r}
par(mfrow=c(2,2))
plot(fit)
```

These diagnostic plots do not reveal any worrying patterns within the residuals, and the data seems nearly normal.


**Comparison of the "Business" variable**

```{r}
business_college <- relevel(as.factor(college$major_category), "Business")
fit2 <- lm(median ~ business_college, college)
round(summary(fit2)$coef, 2)
```

When we compare the `major_category` Business to the other categories, we see a clear difference with some other categories. For example
* Arts
* Computers & Mathematics
* Education
* Engineering
* Humanities & Liberal Arts
* Science
all have negative coefficients, with significantly small p-values (p<0.05)

<br/><br/><br/><br/>





##### Conclusion

This analysis looked at the question of whether there was any difference in income, based on college major.

In order to answer this question, we looked at `median` income, by `major category`.

**Null Hypothesis**

We began with the assumption of *no difference* in income between the major categories. 

The alternative hypothesis, is that there was a difference in mean income between the categories. 


Our regression models show that in general, there is no significant difference in income across major categories. 

We do see **Business** as significantly higher than some other categories, including *Arts, Computers & Mathematics, Education, Engineering, Humanities & Liberal Arts, and Social Science*.

Further analysis would be required to validate and verify if this is a real difference, or just due to chance.
